Co-methylation

### 2015-04-15 Rmarkdown/comethylation.Rmd
Notes following chat with Terry

__TODO: I need to re-write this section (bold). I'm really not yet sure if this result is biology or an artefact. My instinct says artefact but I haven't yet been able to prove this either way to my satisfaction
These plots appear to show that there is a dramatic increase in within-fragment co-methylation at precisely these $IPD$s for paired-end samples from the Lister dataset. However, I am very skeptical as to the biological relevance of these results.

@@

```{r}
# Function to simulate n 2-tuples with d sequencing coverage and marginal 
# probabilities p1, p2
f <- function(n, d, p1, p2) {
  sapply(seq_len(n), function(i, d, p1, p2) {
  z1 <- factor(rbinom(d, 1, p1), levels = c(0, 1))
  z2 <- factor(rbinom(d, 1, p2), levels = c(0, 1))
  x <- xtabs(~ z2 + z1, drop.unused.levels = FALSE)
  log2(x[1, 1] + 0.5) + log2(x[2, 2] + 0.5) - 
    log2(x[1, 2] + 0.5) - log2(x[2, 1] + 0.5)
  }, d = d, p1 = p1, p2 = p2)
}

@@

# Vary p1 with p2 = p1 at 1000x coverage
d <- 1000
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, p)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = p2 = ", p), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)

```

>>>>

__TODO__: I need to re-write this section. I'm really not yet sure if this result is biology or an artefact. My instinct says artefact but I haven't yet been able to prove this either way to my satisfaction
These plots appear to show that there is a dramatic increase in within-fragment co-methylation at precisely these $IPD$s for paired-end samples from the Lister dataset. However, I am very skeptical as to the biological relevance of these results.

@@

```{r}
# Function to simulate n 2-tuples with d sequencing coverage and marginal
# probabilities p1, p2
f <- function(n, d, p1, p2, offset = 0.5) {
  sapply(seq_len(n), function(i, d, p1, p2) {
  z1 <- factor(rbinom(d, 1, p1), levels = c(0, 1))
  z2 <- factor(rbinom(d, 1, p2), levels = c(0, 1))
  x <- xtabs(~ z2 + z1, drop.unused.levels = FALSE)
  log2(x[1, 1] + offset) + log2(x[2, 2] + offset) -
    log2(x[1, 2] + offset) - log2(x[2, 1] + offset)
  }, d = d, p1 = p1, p2 = p2)
}

@@

# Vary p1 with p2 = p1 at 1000x coverage
d <- 1000
n <- 1000
p <- c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99)
x <- lapply(p, function(p) {
  f(n, d, p, p)
})
par(mfrow = c(3, 3))
mapply(function(x, p) {
  hist(x, main = paste0("p1 = p2 = ", p), xlim = c(-10, 10), breaks = 20)
  abline(v = 0, col = "red", lty = 2)
  }, x = x, p = p)
title(paste("d = ", d), outer = TRUE, line = -1)
```

## Following chat with Terry

These are my notes following a Skype conversion with Terry to discuss the issue.

- [ ] Try a higher order Taylor expansion of the log odds-ratio to get e.g., $n^{1}$ or $n^{3/2}$ convergence. The $+\frac{1}{2}$ fudge-factor is motivated by such an expansion and perhaps it is possible to do better by a higher order expansion.
  - [ ] First try showing the $c = \frac{1}{2}$ is optimal.
- [ ] Read texts about this issue
- [ ] Look for research papers on the general issue of dealing with bias in log odds-ratio estimates
- [ ] What if I just increase the fudge-factor to increase the shrinkage/smoothing of the estimates? It'll shrink to zero alright, but what is the theoretical justification?
- [ ] Check out the behaviour of the $\phi$ coefficient, i.e., Pearson's $r$ of the $2 \times 2$ table.
  - Landan et al.: "Note that correlation values are subject to considerable noise when coverage is low or when m1 or m2 are close to 0 or 1."
    - They don't define whether 'noise' means bias, variance, skewness, etc.
